DOM Review Summary

### 2024-03-18 Analysis/ReviewSummary/ReviewSummary.qmd

| Field             | Description                                      | Possible values                      |
|-------------------|--------------------------------------------------|--------------------------------------|
| Journal           | Journal which this article was published in      | *Any*                                |
| Year              | Year of publication                              | *Any*                                |
| Author country    | Nation of the first author's primary affiliation | *ISO 3166 country names*             |
| Author categories | Categories of affiliations for *all* authors     | *Academic, Government, NGO, Private* |

@@

| Objective           | Description                                                                       |
|---------------------|-----------------------------------------------------------------------------------|
| Methods             | Did authors seek to demonstrate or test specific methods for occupancy modelling? |
| Future Predictions  | Did authors make predictions beyond the study period?                             |
| Spatial Predictions | Did authors make predictions at locations where surveys were not conducted?       |
| Hypothesis testing  | Did authors seek to test specific hypotheses of species occupancy dynamics?       |

@@

| Field           | Description                                                                                | Possible values                                                                   |
|-----------------|--------------------------------------------------------------------------------------------|-----------------------------------------------------------------------------------|
| Taxa category   | Category of study taxa; could fit into multiple categories                                 | *Bird, Mammal, Invertebrate, Herptile, Fish, Threatened, Invasive, Other*         |
| Taxa number     | Number of individual species within the taxa for which models were fit                     | *Numeric*                                                                         |
| IUCN status     | IUCN status for the study species; NA if multi-species or if the study species is invasive | *LC, NT, EN, CR, DD, NA*                                                          |
| Threat category | Category of threats to the species; multiple per species permitted                         | *Habitat loss, Climate change, Exploitation, Invasives, Disease, Disaster, Other* |
| Country         | Country/countries where the study took place                                               | *ISO 3166 country names*                                                          |
| Continent       | Continent(s) where the study took place                                                    | *North America, South America, Europe, Asia, Africa, Oceania*                     |
| Realm           | Biogeographic realm where the study took place                                             | *Nearctic, Palearctic, Afrotropical, Indomalayan, Australasian, Neotropical*      |
| Size            | Size of the study area extent in square kilometres                                         | *\<1km, 1-10, 10-100, 100-1000, 10000-100000, 100000-1000000, 1000000+*           |

@@

| Field               | Description                                                              | Possible values                                                                              |
|---------------------|--------------------------------------------------------------------------|----------------------------------------------------------------------------------------------|
| Duration            | Length of study in months                                                | *Numeric*                                                                                    |
| Capture method      | Method of recording detections; multiple categories possible             | *Audiovisual, Search, Trapping, Bioacoustic, Camera, Citizen, Incidental*                    |
| Primary occasions   | Number of primary occasions                                              | *Numeric*                                                                                    |
| Secondary occasions | Number of secondary occasions, presented as a range where not consistent | *Numeric range*                                                                              |
| Secondary type      | Type of secondary occasion                                               | *Revisits, Spatial replicates, Same-visit replicates, multiple observers, time to detection* |
| Type of site        | How a 'site' was defined                                                 | *Point location, Grid cell, Irregular area, Transect*                                        |
| Number of sites     | Number of individual sites                                               | Numeric                                                                                      |
| Sampling method     | How sites were selected                                                  | *Random, Stratified, Grid, Preferential, Available, Exhaustive, Uniform, Transect*           |

@@

| Field         | Description                                               | Possible values                                                                                |
|---------------|-----------------------------------------------------------|------------------------------------------------------------------------------------------------|
| Bayesian      | Whether the models fit were Bayesian                      | *Logical*                                                                                      |
| Prefiltering  | Which, if any, prefiltering method was used               | *Expert, Simple, Other*                                                                        |
| Selection     | Which, if any, selection method was used                  | *Single a priori, Multiple a priori, Exhaustive, Exhaustive by parameter, Forwards, Backwards* |
| Evaluation    | How models were compared and evaluated                    | *AIC, AUC, DiC, Model-averaging, Bootstrapping*                                                |
| Multi-species | Whether this was a multi-species model                    | *Logical*                                                                                      |
| Connectivity  | Whether site connectivity was incorporated into the model | *Logical*                                                                                      |

>>>>

+-------------------+--------------------------------------------------+--------------------------------------+
| Field             | Description                                      | Possible values                      |
+===================+==================================================+======================================+
| Journal           | Journal which this article was published in      | *Any*                                |
+-------------------+--------------------------------------------------+--------------------------------------+
| Year              | Year of publication                              | *Any*                                |
+-------------------+--------------------------------------------------+--------------------------------------+
| Author country    | Nation of the first author's primary affiliation | *ISO 3166 country names*             |
+-------------------+--------------------------------------------------+--------------------------------------+
| Author categories | Categories of affiliations for *all* authors     | *Academic, Government, NGO, Private* |
+-------------------+--------------------------------------------------+--------------------------------------+

@@

+---------------------+-----------------------------------------------------------------------------------+
| Objective           | Description                                                                       |
+=====================+===================================================================================+
| Methods             | Did authors seek to demonstrate or test specific methods for occupancy modelling? |
+---------------------+-----------------------------------------------------------------------------------+
| Future Predictions  | Did authors make predictions beyond the study period?                             |
+---------------------+-----------------------------------------------------------------------------------+
| Spatial Predictions | Did authors make predictions at locations where surveys were not conducted?       |
+---------------------+-----------------------------------------------------------------------------------+
| Hypothesis testing  | Did authors seek to test specific hypotheses of species occupancy dynamics?       |
+---------------------+-----------------------------------------------------------------------------------+

@@

+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Field           | Description                                                                                | Possible values                                                                   |
+=================+============================================================================================+===================================================================================+
| Taxa category   | Category of study taxa; could fit into multiple categories                                 | *Bird, Mammal, Invertebrate, Herptile, Fish, Threatened, Invasive, Other*         |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Taxa number     | Number of individual species within the taxa for which models were fit                     | *Numeric*                                                                         |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| IUCN status     | IUCN status for the study species; NA if multi-species or if the study species is invasive | *LC, NT, EN, CR, DD, NA*                                                          |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Threat category | Category of threats to the species; multiple per species permitted                         | *Habitat loss, Climate change, Exploitation, Invasives, Disease, Disaster, Other* |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Country         | Country/countries where the study took place                                               | *ISO 3166 country names*                                                          |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Continent       | Continent(s) where the study took place                                                    | *North America, South America, Europe, Asia, Africa, Oceania*                     |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Realm           | Biogeographic realm where the study took place                                             | *Nearctic, Palearctic, Afrotropical, Indomalayan, Australasian, Neotropical*      |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
| Size            | Size of the study area extent in square kilometres                                         | *\<1km, 1-10, 10-100, 100-1000, 10000-100000, 100000-1000000, 1000000+*           |
+-----------------+--------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+

@@

+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Field               | Description                                                              | Possible values                                                                              |
+=====================+==========================================================================+==============================================================================================+
| Duration            | Length of study in months                                                | *Numeric*                                                                                    |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Capture method      | Method of recording detections; multiple categories possible             | *Audiovisual, Search, Trapping, Bioacoustic, Camera, Citizen, Incidental*                    |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Primary occasions   | Number of primary occasions                                              | *Numeric*                                                                                    |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Secondary occasions | Number of secondary occasions, presented as a range where not consistent | *Numeric range*                                                                              |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Secondary type      | Type of secondary occasion                                               | *Revisits, Spatial replicates, Same-visit replicates, multiple observers, time to detection* |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Type of site        | How a 'site' was defined                                                 | *Point location, Grid cell, Irregular area, Transect*                                        |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Number of sites     | Number of individual sites                                               | Numeric                                                                                      |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+
| Sampling method     | How sites were selected                                                  | *Random, Stratified, Grid, Preferential, Available, Exhaustive, Uniform, Transect*           |
+---------------------+--------------------------------------------------------------------------+----------------------------------------------------------------------------------------------+

@@

+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Field         | Description                                               | Possible values                                                                                |
+===============+===========================================================+================================================================================================+
| Bayesian      | Whether the models fit were Bayesian                      | *Logical*                                                                                      |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Prefiltering  | Which, if any, prefiltering method was used               | *Expert, Simple, Other*                                                                        |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Selection     | Which, if any, selection method was used                  | *Single a priori, Multiple a priori, Exhaustive, Exhaustive by parameter, Forwards, Backwards* |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Evaluation    | How models were compared and evaluated                    | *AIC, AUC, DiC, Model-averaging, Bootstrapping*                                                |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Multi-species | Whether this was a multi-species model                    | *Logical*                                                                                      |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+
| Connectivity  | Whether site connectivity was incorporated into the model | *Logical*                                                                                      |
+---------------+-----------------------------------------------------------+------------------------------------------------------------------------------------------------+

```{r modelling_clean}
#| echo: false

modelling_clean <- review_Modelling |>
  filter(Status == "Complete")
```

## Selection methods

```{r modelling_selection}
#| echo: false

modelling_prefilter <- modelling_clean |>
  mutate(Total = length(unique(ID))) |>
  summarise(Articles = n(), 
            .by = c(Prefiltering, Total)) |>
  rename(Method = Prefiltering) |>
  mutate(Class = "Prefilter")

modelling_clean |> 
  mutate(Total = length(unique(ID))) |>
  summarise(Articles = n(), 
            .by = c(Selection, Total)) |>
  rename(Method = Selection) |>
  mutate(Class = "Select") |>
  bind_rows(modelling_prefilter) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  filter(!Method %in% c("None", "APriori_m, Stepwise", "Other")) |>
  mutate(Method = case_when(Method %in% c("Apriori_s", "APriori_s") ~ "A priori (Single)",
                            Method == "APriori_m" ~ "A priori (Multiple)",
                            Method == "Backwards" ~ "Stepwise (Backwards)",
                            Method == "Forwards" ~ "Stepwise (Forwards)",
                            Method == "Exhaustive_all" ~ "Exhaustive (True)",
                            Method == "Exhaustive_p" ~ "Exhaustive (By parameter)",
                            .default = Method)) |>
  summarise(Percentage = sum(Percentage),
            .by = c(Method, Class)) |>
  arrange(Class, Percentage) |>
  mutate(Method = fct_inorder(Method)) |>

  ggplot() +
  geom_col(aes(x = Percentage, y = Method, fill = Class)) +
  scale_fill_manual(limits = c("Prefilter", "Select"),
                    values = c("darkorange3", "darkslategrey")) +
  labs(title = "Covariate selection method frequency", 
       x = "Percentage of articles", y = "") +
  theme(legend.position = "bottom")
```

## Model selection trends

```{r Modelling_trends}
#| echo: false
#| warning: false

modelling_clean |>
  select(ID, Selection) |>
  mutate(ID = as.numeric(ID)) |>
  left_join(select(studyDetails_clean, ID, YearStrata),
            by = "ID") |>
  filter(!is.na(YearStrata)) |>
  mutate(Selection = case_when(Selection %in% c("APriori_s",
                                                "Apriori_s") ~ "A priori (Single)",
                               Selection == "APriori_m" ~ "A priori (Multiple)",

                               .default = "Other")) |>
  mutate(Total = n(), .by = YearStrata) |>
  summarise(Articles = n(), .by = c(YearStrata, Selection, Total)) |>
  mutate(Proportion = Articles/Total) |>

  ggplot() +
  geom_col(aes(x = YearStrata, y = Articles, fill = Selection)) +
  scale_fill_manual(values = c("goldenrod", "darkorange3", "darkorchid4")) +
  theme(legend.position = "bottom") +
  labs(title = "Covariate selection trends",
       x = "Year Strata", 
       y = "Number of articles")

## Frequentist vs Bayesian

```{r modelling_frequentist}
#| echo: false
#| warning: false

modelling_clean |>
  mutate(Class = case_when(Bayesian == "Y" ~ "Bayesian",
                           Bayesian == "N" ~ "Frequentist")) |>
  mutate(ID = as.numeric(ID)) |>
  left_join(select(studyDetails_clean, ID, YearStrata),
            by = "ID") |>
  filter(!is.na(YearStrata),
         !is.na(Class)) |>
  summarise(Articles = n(),
            .by = c(YearStrata, Class)) |>

  ggplot() +
  geom_col(aes(x = YearStrata, y = Articles, fill = Class)) +
  scale_fill_manual(values = c("cyan4", "darkorchid3")) +
  theme(legend.position = "bottom") +
  labs(title = "Model type trends",
       x = "Year Strata", 
       y = "Number of articles")
